---
title: "01_Pipeline_DEA"
author: "Chang Lu"
date: "5/11/2020"
---



## Introduction

This is the pipeline of Differential Gene Analysis of Gender Study in CVD on combined data (GSE 9820 and CTMM cohort).


## library packages
```{r}
library(pracma)
library(limma)
library(xlsx)
library(BatchQC)
library(factoextra)
library(lumi)
library(dplyr)
```



```{r}
##loading GSE 9820 and CTMM cohort data
load('./data/Dataforlimma30demo.RData') 
# clinforlimma :clinical informations of patients: sex, ages, batch and indication
# exprforlimma :gene expression data

```



```{r}
## gene ID to gene hgnc_symbol
require(lumiHumanIDMapping)

mappingInfo <- getNuIDMappingInfo(lib.mapping='lumiHumanIDMapping') %>% 
  dplyr::rename(genes = 'EntrezID') %>% distinct(genes, .keep_all= TRUE)

geneInfo = data.frame(genes = rownames(exprforlimma), row.names = rownames(exprforlimma))  
geneInfo = geneInfo %>% 
  left_join(mappingInfo %>% dplyr::select(genes,Symbol),by='genes') %>%
  dplyr::rename(hgnc_symbol = 'Symbol')
```


# limma (add batches as coviarates)
```{r}
TS <- paste(clinforlimma$sex,clinforlimma$indication, sep=".")
TS <- factor(TS, levels=unique(TS))

batch <- as.factor(clinforlimma$batch)
design2 <- model.matrix(~0 + TS + batch + clinforlimma$age)
colnames(design2)[1:4] <- c("male.CVD","male.healthy","female.CVD" ,"female.healthy")
colnames(design2)[ncol(design2)] <- c("age")

fit.limma <- lmFit(exprforlimma, design2)
cont.matrix <- makeContrasts(
  CVDvsHealthyinMale = male.CVD - male.healthy,
  CVDvsHealthyinFemale = female.CVD - female.healthy,
  CVDMalevsFemale = male.CVD - female.CVD,
  HealthyMalevsFemale = male.healthy - female.healthy,
  levels = design2)
fit.limma2 <- contrasts.fit(fit.limma, cont.matrix)
fit.limma2 <- eBayes(fit.limma2)

# plotMD(fit.limma2,column=3)
# volcanoplot(fit.limma2,coef=4,highlight=2)
```




```{r}
# draw venn diagram
results <- decideTests(fit.limma2, adjust.method = "fdr", p.value = 0.05)
jpeg('./results/venn_limma_correctage_pv=0.05_fdr.jpg',width = 800, height = 600)
vennDiagram(results)
dev.off()

```




```{r}
# DEA results
male_DEG_limma = topTable(fit.limma2, coef="CVDvsHealthyinMale",   number = 20000, p.value = 1, adjust.method="fdr")
female_DEG_limma = topTable(fit.limma2, coef="CVDvsHealthyinFemale", number = 20000,p.value = 1, adjust.method="fdr")
CVDMF_DEG_limma = topTable(fit.limma2, coef="CVDMalevsFemale",   number = 20000, p.value = 1, adjust.method="fdr")
HealthMF_DEG_limma = topTable(fit.limma2, coef="HealthyMalevsFemale", number = 20000,p.value = 1, adjust.method="fdr")
```




```{r}
# draw volcano plot
volcanoplot(fit.limma2, coef = 'CVDvsHealthyinMale')
title('CVD vs Healthy on Male')
volcanoplot(fit.limma2, coef = "CVDvsHealthyinFemale")
title('CVD vs Healthy on female')
```


```{r}
#save data 
save(fit.limma2, male_DEG_limma, female_DEG_limma, CVDMF_DEG_limma, HealthMF_DEG_limma, clinforlimma ,exprforlimma, geneInfo, file = './data/batch-limma-CVD-Control.RData')
```








Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
