---
title: "03_Pipeline_PROGENY"
output: html_notebook
---

```{r}
library(progeny)
library(gridExtra)
library(ggrepel)
library(reshape)
library(dplyr)
library(tibble)
library(tidyr)
library(viper)
source('./fun/progeny_function.R')
library(ComplexHeatmap)
library(ggnewscale)
```




# download the raw progeny weight matrix
```{r}
prog_matrix_full = get("model_human_full", envir = .GlobalEnv)
prog_w = reshape(prog_matrix_full[c("gene", "pathway",'weight')], idvar = "gene", timevar = "pathway", direction = "wide")
colnames(prog_w) = stringr::str_remove(colnames(prog_w),'weight.')

## calculating the adjusted pvalue across all weights (all pathways)
prog_pv = reshape(prog_matrix_full[c("gene", "pathway",'p.value')], idvar = "gene", timevar = "pathway", direction = "wide")
colnames(prog_pv) = stringr::str_remove(colnames(prog_pv),'p.value.')
prog_matrix_full$adj.pv = p.adjust(prog_matrix_full$p.value, method = 'fdr', n = nrow(prog_matrix_full))
prog_adjpv2 = reshape(prog_matrix_full[c("gene", "pathway",'adj.pv')], idvar = "gene", timevar = "pathway", direction = "wide")
colnames(prog_adjpv2) = stringr::str_remove(colnames(prog_adjpv2),'adj.pv.')


## calculating the adjusted pvalue for each pathway
prog_adjpv = apply(prog_pv[,2:15], 2, function(x) {
  adjp = matrix(nrow = length(x), ncol = 1)
  xi = x[!is.na(x)]
  adjp[!is.na(x)] = p.adjust(xi, method = 'fdr', n = length(xi))
  return(adjp)
})
prog_adjpv = data.frame(prog_adjpv)
prog_adjpv =  cbind(prog_pv$gene,prog_adjpv)
colnames(prog_adjpv)[1] = 'gene'

```



## distribution of gene weights per pathway in 'model_human_full'
```{r}
jpeg(paste0("./results/progenyweight.jpg"), height = 800, width = 1200)
ggplot(prog_matrix_full, aes(weight)) + 
  geom_density()+
  facet_wrap(~pathway,ncol=3)+scale_x_continuous(limits = c(-10, 10))+
  theme(axis.text.y = element_text(face = "bold", size = 15),
        axis.text.x = element_text(face = "bold", size = 15),
        axis.title=element_text(size=20,face="bold"),
        strip.text = element_text(size = 15),
        legend.text = element_text(size = 15))+
  xlab("weight")+ylab("density")
dev.off()


jpeg(paste0("./progenyweightboxplot.jpg"), height = 600, width = 1000)
ggplot(prog_matrix_full,aes(x= pathway, y = weight)) + 
  geom_boxplot()+
  theme(axis.text.y = element_text(face = "bold", size = 15),
        axis.text.x = element_text(face = "bold", size = 15),
        axis.title=element_text(size=20,face="bold"),
        strip.text = element_text(size = 15),
        legend.text = element_text(size = 15))+
  xlab("pathway")+ylab("weight")
dev.off()

```


```{r}
## load female's and male's GES data (from '02_Pipeline_EA.Rmd')
load('./data/GES_ma_fe.RData')
load('./data/batch-limma-CVD-Control.RData')
finalclin = clinforlimma
finalgenemat = exprforlimma


```


```{r}
## extract the genes in Our dataset and check the distributions

new_prog_matrix_full = prog_matrix_full[prog_matrix_full$gene %in% GES_fe_HGNC$ID,] 

jpeg(paste0("./results/subprogenyweight.jpg"), height = 800, width = 1200)
ggplot(new_prog_matrix_full, aes(weight)) + 
  geom_density()+
  facet_wrap(~pathway,ncol=3)+scale_x_continuous(limits = c(-10, 10))+
  theme(axis.text.y = element_text(face = "bold", size = 15),
        axis.text.x = element_text(face = "bold", size = 15),
        axis.title=element_text(size=20,face="bold"),
        strip.text = element_text(size = 15),
        legend.text = element_text(size = 15))+
  xlab("weight")+ylab("density")
dev.off()

jpeg(paste0("./results/subprogenyweightboxplot.jpg"), height = 600, width = 1000)
ggplot(new_prog_matrix_full,aes(x= pathway, y = weight)) + 
  geom_boxplot()+
  theme(axis.text.y = element_text(face = "bold", size = 15),
        axis.text.x = element_text(face = "bold", size = 15),
        axis.title=element_text(size=20,face="bold"),
        strip.text = element_text(size = 15),
        legend.text = element_text(size = 15))+
  xlab("pathway")+ylab("weight")
dev.off()





```



## check the NES stability based on adjusted pvalue (prog_adjpv2) and top n genes
## gene permulation (you can also choose sample permutation)
```{r}
source('./fun/progeny_function.R')
toplist = seq(from = 10, to = 1000, by = 200)
flagname = 'prog_adjpv2 topn genePerm'

progScorelisttopadj_zs = sapply(toplist,function(x) getprogenyScore(pvmat = prog_adjpv2, 
                                                                    top = x, 
                                                                    zscore = TRUE,
                                                                    perm.by = 'gene')) # or 'sample'

prog_mat_p_zslist =lapply(seq(from = 1, to = length(progScorelisttopadj_zs)-1, by = 2), function(x) progScorelisttopadj_zs[[x]])
progScorelist_zs = sapply(seq(from = 2, to = length(progScorelisttopadj_zs), by = 2), function(x) progScorelisttopadj_zs[[x]])


progScorelisttopadj = sapply(toplist,function(x) getprogenyScore(pvmat = prog_adjpv2, 
                                                                 top = x, 
                                                                 zscore = FALSE,
                                                                 perm.by = 'gene'))
prog_mat_p_list = lapply(seq(from = 1, to = length(progScorelisttopadj)-1, by = 2), function(x) progScorelisttopadj[[x]])
progScorelist = sapply(seq(from = 2, to = length(progScorelisttopadj), by = 2), function(x) progScorelisttopadj[[x]])

## visualize the stability of NES by line chart
linePlotVis(progScorelist, filename = paste0('./results/',flagname,'.jpg'), toplist, xlabel= 'top n')
linePlotVis(progScorelist=progScorelist_zs, filename = paste0('./results/',flagname,'_zscore.jpg'), toplist=toplist, xlabel= 'top n')

```




## check the NES stability based on adjusted pvalue (prog_adjpv2) with increase of 'adjusted pvalue cutoff'
## sample permulation
```{r}
source('./fun/progeny_function.R')
source('./fun/viper_function.R')
toplist = seq(from = 1, to = 4, by = 0.1)
flagname = 'prog_adjpv2 pcutoff samplePerm'

# sample permutation using limma 
tictoc::tic()
dnullma <- viperNulllimma(expset=exprs_hgnc_ma, clin=clin_ma, per = 100, repos=TRUE)
tictoc::toc()
tictoc::tic()
dnullfe <- viperNulllimma(expset=exprs_hgnc_fe, clin=clin_fe, per = 100, repos=TRUE)
tictoc::toc()


progadjPVlist_permsam = sapply(toplist,function(x) getprogenyScorePV(pvmat = prog_adjpv2, 
                                                                     logpv=x,
                                                                     zscore = FALSE,
                                                                     perm.by = 'sample',
                                                                     dnullma = dnullma,
                                                                     dnullfe = dnullfe)) #or 'gene'
progadjPVlist_permsam_zs = sapply(toplist,function(x) getprogenyScorePV(pvmat = prog_adjpv2, 
                                                                     logpv=x,
                                                                     zscore = TRUE,
                                                                     perm.by = 'sample',                                                                                                                             dnullma = dnullma,
                                                                     dnullfe = dnullfe)) #or 'gene'

prog_mat_p_list = lapply(seq(from = 1, to = length(progadjPVlist_permsam)-1, by = 2), function(x) progadjPVlist_permsam[[x]])
progScorelist = sapply(seq(from = 2, to = length(progadjPVlist_permsam), by = 2), function(x) progadjPVlist_permsam[[x]])
prog_mat_p_zslist = lapply(seq(from = 1, to = length(progadjPVlist_permsam_zs)-1, by = 2), function(x) progadjPVlist_permsam_zs[[x]])
progScorelist_zs = sapply(seq(from = 2, to = length(progadjPVlist_permsam_zs), by = 2), function(x) progadjPVlist_permsam_zs[[x]])


## visualize the stability of NES by line chart
linePlotVis(progScorelist=progScorelist, filename = paste0('./results/',flagname,'.jpg'), toplist, xlabel= '-log10(adj.pv)')
linePlotVis(progScorelist=progScorelist_zs, filename = paste0('./results/',flagname,'_zscore.jpg'), toplist=toplist, xlabel= '-log10(adj.pv)')

```


## save progeny results for CARNIVAL
```{r}
progScorelist_SamPerm = sapply(seq(from = 2, to = length(progadjPVlist_permsam), by = 2), function(x) progadjPVlist_permsam[[x]])
saveRDS(progScorelist_SamPerm, file = './data/progeny4carnival_sampleperm.rds')

```



# all steps below were based on the results of 'prog_adjpv2 pcutoff samplePerm'
# prog_mat_p_zslist data tube similalrity based on 'prog_adjpv2 pcutoff samplePerm'
```{r}
top = 1:4
for (i in top) {
  idx = which(toplist == i)
  prog_mat_p = prog_mat_p_zslist[[idx]]
  prog_mat_p_cor = cor(prog_mat_p[,2:15],method = "spearman")
  jpeg(paste0('./results/corr_14pathway_zcoff_spearman_pv',i,'.jpg'), height = 1100, width = 1000, res = 180)
  p = draw(Heatmap(prog_mat_p_cor, 
                   name = "Correlation",
                   heatmap_legend_param = list(direction = "horizontal")),
           heatmap_legend_side = "bottom")
  print(p)
  dev.off()
}

jpeg('./results/corr_14pathway_zcoff_spearman_pv0.jpg', height = 1100, width = 1000, res = 180)
prog_mat_p_cor = cor(prog_w[,2:15],use = "na.or.complete",method = "spearman")
draw(Heatmap(prog_mat_p_cor, 
             name = "Correlation",
             heatmap_legend_param = list(direction = "horizontal")),
     heatmap_legend_side = "bottom")
dev.off()

```


# draw a scatter plot to visualize female and male specific TF regluons' avtivity at when adj.pv = 3
```{r}
pcutoff = 3
idx = which(toplist == pcutoff)

# # norm
# progenyScore = sapply(seq(3*idx-1,3*idx), function(x) as.numeric(progScorelist[[x]]))  %>% 
#   as.data.frame() %>% 
#   add_column(Pathway = progScorelist[[1]])
# flagname2 = flagname
# colnames(progenyScore) <- c('male','female','Pathway')
# prog_mat_p = prog_mat_p_list[[idx]]

# # opposite pw
# cyl = data.frame(rep('Others',nrow(progenyScore)) ,row.names = progenyScore$Pathway)
# cyl[abs(progenyScore$female)>0.9 & abs(progenyScore$female-progenyScore$male)>0.5,]<- "Female specific Pathway"
# cyl[abs(progenyScore$male)>0.9& abs(progenyScore$female-progenyScore$male)>0.5,]<- "Male specific Pathway"
# cyl[abs(progenyScore$female)>0.9 & abs(progenyScore$male)>0.9,] <- "Shared significant Pathway"
# progenyScore$labels <- NA
# progenyScore$labels <- cyl[,1]


# zscore
progenyScore = sapply(seq(3*idx-1,3*idx), function(x) as.numeric(progScorelist_zs[[x]]))  %>% 
  as.data.frame() %>% 
  add_column(Pathway = progScorelist_zs[[1]])
flagname2 = paste(flagname,'zscore')
colnames(progenyScore) <- c('male','female','Pathway')
prog_mat_p = prog_mat_p_list[[idx]]

# female/male specific TF
cyl = data.frame(rep('Others',nrow(progenyScore)) ,row.names = progenyScore$Pathway)
cyl[abs(progenyScore$female)>1.96 & abs(progenyScore$female-progenyScore$male)>2,]<- "Female specific Pathway"
cyl[abs(progenyScore$male)>1.96& abs(progenyScore$female-progenyScore$male)>2,]<- "Male specific Pathway"
cyl[abs((progenyScore$male)>1.96 | abs(progenyScore$female)>1.96)& abs(progenyScore$female-progenyScore$male)<1,] <-"Shared significant Pathway"
cyl[abs(progenyScore$female)>1.96 & abs(progenyScore$male)>1.96,] <- "Shared significant Pathway"
progenyScore$labels <- NA
progenyScore$labels <- cyl[,1]


# scatter
jpeg(paste0("./results/tstat_ProgenyscorePV_FvsM_TOP",as.character(10^(-pcutoff)), flagname2, ".jpg"), 
     height = 800, width = 1200, res = 180)
ggplot(data = progenyScore, aes(x = male, y = female, color = female-male)) + 
  geom_point(size = 3)+
  scale_color_gradient2(midpoint = 0, low = "darkblue", mid = "white",high = "darkred", space = "Lab" )+
  new_scale('color')+
  geom_text_repel(aes(label = Pathway, color = labels),fontface = "bold")+
  scale_colour_manual(values = c("Female specific Pathway" = "red",
                                 "Male specific Pathway" = "blue",
                                 "Shared significant Pathway" = "black",
                                 "Others" = "grey"))+
  xlab('NES on Male')+ylab('NES on Female')+
  ggtitle("Pathway Activity (Female vs Male)") 
dev.off()


```



# draw a vocalno plot to visualize PI3K on female and male GES
```{r}
i = "PI3K"
intersectgenes <- intersect(prog_w$gene[!is.na(prog_w[,i])],GES_fe_HGNC$ID)
prog_wi =  prog_w[, c('gene',i)] %>% dplyr::filter(gene %in% intersectgenes) %>% dplyr::rename(ID = 'gene' )
tmat_fe = merge(prog_wi, GES_fe_HGNC[GES_fe_HGNC$ID %in% intersectgenes,], by = 'ID')
tmat_ma = merge(prog_wi, GES_ma_HGNC[GES_ma_HGNC$ID %in% intersectgenes,], by = 'ID')
colnames(tmat_fe)[2] = colnames(tmat_ma)[2] = 'Progeny.weight'
ProgenyvocalnoPlot(tmat= tmat_fe, 
                   filename = paste0('./results/',i,' female ProgenyvocalnoPlot.jpg'), 
                   xname = 'logFC',
                   yname = "adj.P.Val",
                   labelby = "adj.P.Val",
                   weightname = 'Progeny.weight',
                   top = 10)
ProgenyvocalnoPlot(tmat= tmat_ma, 
                   filename = paste0('./results/',i,' male ProgenyvocalnoPlot.jpg'), 
                   xname = 'logFC',
                   yname = "adj.P.Val",
                   labelby = "adj.P.Val",
                   weightname = 'Progeny.weight',
                   top = 10) 


```



## scatter for visualizing female's and male's logFC and the most important genes that make the greatest contributions to pathway activity calculation
```{r}
# i = "NFkB"
# i = "PI3K"    
# i = "TGFb"     
# i = "TNFa"  
# i = "EGFR" 
# i = "JAK-STAT"

pathname = colnames(prog_w)[2:15]
for(i in pathname){
  
 intersectgenes <- intersect(prog_w$gene[!is.na(prog_w[,i])],GES_fe_HGNC$ID)
 prog_wi =  prog_w[, c('gene',i)] %>% dplyr::filter(gene %in% intersectgenes) 
 colnames(prog_wi) = c("ID", "Progeny.weight")
 tmat = merge(prog_wi,GES_fe_HGNC[GES_fe_HGNC$ID %in% intersectgenes,],by = 'ID')
 tmat = merge(tmat,GES_ma_HGNC[GES_ma_HGNC$ID %in% intersectgenes,],by = 'ID',suffixes = c(".female",".male"))

 
 tmat <- tmat %>% dplyr::mutate(contribute.male = Progeny.weight*t.male, contribute.female = Progeny.weight*t.female)
 ProgenyScatterPlot(tmat= tmat, 
                    filename = paste0('./results/',i,' ProgenyScatter-fe-logfc.jpg'), 
                    xname = 'male',
                    yname = "female",
                    valuename = 'logFC',
                    weightname = 'Progeny.weight',
                    labelby = 'contribute.female',
                    top = 10) 
 
 ProgenyScatterPlot(tmat= tmat, 
                    filename = paste0('./results/',i,' ProgenyScatter-ma-logfc.jpg'), 
                    xname = 'male',
                    yname = "female",
                    valuename = 'logFC',
                    weightname = 'Progeny.weight',
                    labelby = 'contribute.male',
                    top = 10) 
 
 
}

```
























Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
