---
title: "04_Pipeline_tfactivity"
output: html_notebook
---



```{r}
library(gridExtra)
library(ggrepel)
library(reshape)
library(dplyr)
library(tibble)
library(tidyr)
library(viper)
source('./fun/viper_function.R')
library(ComplexHeatmap)
library(dorothea)
library(egg)
library(xlsx)
library(ggpubr)
source('./fun/multiORA.R')
```



```{r}
#load Dorothea Regulons
data(dorothea_hs, package = "dorothea")
regulons <- dorothea_hs %>%
  dplyr::filter(confidence %in% c("A", "B","C"))
regulons2 <- df2regulon(regulons)

regulonsALL <- dorothea_hs %>%
  dplyr::filter(confidence %in% c("A", "B","C","D","E"))
regulonsALL2 <- df2regulon(regulonsALL)

```


```{r}
## load our GES data (from '02_Pipeline_EA.Rmd')
load('./data/GES_ma_fe.RData')
load('./data/batch-limma-CVD-Control.RData')
finalclin = clinforlimma
finalgenemat = exprforlimma


```


## calculate enrichment score using msviper  (gene permutation)
```{r}

# tictoc::tic()
# dnull <- viperNullgene(expset=GES_ma_mat, per= 20000)
# tictoc::toc()
# TFactivity_ma <- msviper(GES_ma_mat, regulons2, dnull, minsize = 5, cores = 1)
# tf_activities_ma = data.frame(NES = TFactivity_ma$es$nes, 
#                               PV = TFactivity_ma$es$p.value,
#                               TF = names(TFactivity_ma$es$nes))
# tf_activities_ma$p.adj <- p.adjust(tf_activities_ma$PV, method = 'fdr', n = nrow(tf_activities_ma))
# 
# 
# tictoc::tic()
# dnull <- viperNullgene(expset=GES_fe_mat, per= 20000)
# tictoc::toc()
# TFactivity_fe <- msviper(GES_fe_mat, regulons2, dnull, minsize = 5, cores = 1)
# tf_activities_fe = data.frame(NES = TFactivity_fe$es$nes, 
#                               PV = TFactivity_fe$es$p.value,
#                               TF = names(TFactivity_fe$es$nes))
# tf_activities_fe$p.adj <- p.adjust(tf_activities_fe$PV, method = 'fdr', n = nrow(tf_activities_fe))
# 
# rm(dnull)
# TF_activity_all <- merge(tf_activities_ma, tf_activities_fe, by='TF', all=TRUE)
# colnames(TF_activity_all) <- c('TF', 'NES.male','PV.male','p.adj.male','NES.female','PV.female','p.adj.female')
# flagname = 'GenePerm'

```




## calculate enrichment score using msviper  (sample permutation)
```{r}
tictoc::tic()
dnullma <- viperNulllimma(expset=exprs_hgnc_ma, clin=clin_ma, per = 100, repos=TRUE)
tictoc::toc()
tictoc::tic()
dnullfe <- viperNulllimma(expset=exprs_hgnc_fe, clin=clin_fe, per = 100, repos=TRUE)
tictoc::toc()


TFactivity_ma <- msviper(GES_ma_mat, regulons2, dnullma, minsize = 5, cores = 1)
# plot(TFactivity_ma, mrs = 30, cex=1, pval = 0.00)
tf_activities_ma_sampperm = data.frame(NES = TFactivity_ma$es$nes, 
                              PV = TFactivity_ma$es$p.value,
                              TF = names(TFactivity_ma$es$nes))
tf_activities_ma_sampperm$p.adj <- p.adjust(tf_activities_ma_sampperm$PV, method = 'fdr', n = nrow(tf_activities_ma_sampperm))


TFactivity_fe <- msviper(GES_fe_mat, regulons2, dnullfe, minsize = 5, cores = 1)
tf_activities_fe_sampperm = data.frame(NES = TFactivity_fe$es$nes, 
                              PV = TFactivity_fe$es$p.value,
                              TF = names(TFactivity_fe$es$nes))
tf_activities_fe_sampperm$p.adj <- p.adjust(tf_activities_fe_sampperm$PV, method = 'fdr', n = nrow(tf_activities_fe_sampperm))


TF_activity_all <- merge(tf_activities_ma_sampperm, tf_activities_fe_sampperm, by='TF', all=TRUE)
colnames(TF_activity_all) <- c('TF', 'NES.male','PV.male','p.adj.male','NES.female','PV.female','p.adj.female')
flagname = 'samplePerm'
```






# TF activity visualization by 2d plot (male vs female)
```{r}
# TF activity visualization by 2d plot
cyl = data.frame(rep(NA,nrow(TF_activity_all)) ,row.names = TF_activity_all$TF)
cyl[abs(TF_activity_all$NES.female)>1.96 & abs(TF_activity_all$NES.male)<1.96 & abs(TF_activity_all$NES.female-TF_activity_all$NES.male)>2 ,] <- 'Female specific TF' 
cyl[abs(TF_activity_all$NES.male)>1.96 & abs(TF_activity_all$NES.female)<1.96 & abs(TF_activity_all$NES.female-TF_activity_all$NES.male)>2 ,] <- 'Male specific TF' 
cyl[abs((TF_activity_all$NES.male)>1.96 | abs(TF_activity_all$NES.female)>1.96)& abs(TF_activity_all$NES.female-TF_activity_all$NES.male)<1,] <- 'Shared TF'
cyl[abs(TF_activity_all$NES.male)>1.96 & abs(TF_activity_all$NES.female)>1.96,] <- 'Shared TF'
TF_activity_all$labels <- NA
TF_activity_all$labels <- factor(cyl[,1])


jpeg(paste0("./results/TF_FvsM",flagname,".jpg"), height = 1000, width = 1300, res = 150)
ggplot(data = TF_activity_all, aes(x = NES.male, y = NES.female, color = NES.female-NES.male)) +
  geom_point(size = 3)+
  scale_color_gradient2(midpoint = 0, 
                        low = "darkblue", mid = "white",high = "darkred", space = "Lab" )+
  new_scale('color')+
  geom_label_repel(aes(label = TF, color = labels),fontface = "bold", 
                  data = dplyr::filter(TF_activity_all, !is.na(labels)))+
  scale_colour_manual(values = c("Female specific TF" = "red", "Male specific TF" = "blue", "Shared TF" = "black"))+
  # ggtitle("TF activities (Female vs male)")+
  labs(x = "TF activity males (NES value)",y="TF activity females (NES value)")+
  theme(
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 15))
dev.off()



```




## 2D plot of a TF (female vs Male)
```{r}
GES_HGNC_ALL <- merge(GES_ma_HGNC, GES_fe_HGNC, by='ID', all=TRUE, suffixes = c(".ma",".fe"))
tfname = c("IRF3","HNF1A")
i=1

p <- TF2dplot(GES_HGNC_ALL,  tfname[i], axis = 't', label.by='P.Value', cutoff = 0.05)
jpeg(paste0("./results/",tfname[i]," TF.jpg"), width = 1900, height = 1900, res = 170)
print(p[[i]])
dev.off()


```



## SAVE DATA FOR CARNICAL
```{r}

mergeTF <- function(tfactma,tfactfe){
TF_activity_all <- merge(tfactma, tfactfe, by='TF', all=TRUE)
colnames(TF_activity_all) <- c('TF', 'NES.male','PV.male','p.adj.male','NES.female','PV.female','p.adj.female')
cyl = data.frame(rep(NA,nrow(TF_activity_all)) ,row.names = TF_activity_all$TF)
cyl[abs(TF_activity_all$NES.female)>1.96 & abs(TF_activity_all$NES.male)<1.96 & abs(TF_activity_all$NES.female-TF_activity_all$NES.male)>2 ,] <- 'Female specific TF' 
cyl[abs(TF_activity_all$NES.male)>1.96 & abs(TF_activity_all$NES.female)<1.96 & abs(TF_activity_all$NES.female-TF_activity_all$NES.male)>2 ,] <- 'Male specific TF' 
cyl[abs((TF_activity_all$NES.male)>1.96 | abs(TF_activity_all$NES.female)>1.96)& abs(TF_activity_all$NES.female-TF_activity_all$NES.male)<1,] <- 'Shared TF'
cyl[abs(TF_activity_all$NES.male)>1.96 & abs(TF_activity_all$NES.female)>1.96,] <- 'Shared TF'
TF_activity_all$labels <- NA
TF_activity_all$labels <- factor(cyl[,1])
return(TF_activity_all)
}

TFactivity_Carnival_sampPerm <- mergeTF(tf_activities_ma_sampperm, tf_activities_fe_sampperm)

saveRDS(TFactivity_Carnival_sampPerm, file = './data/TF4Carnival_samplePerm.rds')

```




#ORA
```{r}
GOFILE = './referencefile/c5.go.v7.4.symbols.gmt'
pathways = "./referencefile/c2.cp.v7.2.symbols.gmt"
bgtf = unique(regulons$tf)
bgta <- regulons$target[TF_activity_all$TF %in% bgtf]
bg <-  unique(c(TF_activity_all$TF,bgta))


CHOICE = expand.grid(reg = c('up','dn'),
                     database = c('pathway','GO'))
CHOICE = sapply(CHOICE,as.character)
listname <- apply(CHOICE, 1 , function(x) paste(x, collapse = "_"))


TF_male_geneset = TF_activity_all[TF_activity_all$labels %in% 'Male specific TF',]
TF_female_geneset = TF_activity_all[TF_activity_all$labels %in% 'Female specific TF',]
TF_shared_geneset = TF_activity_all[TF_activity_all$labels %in% 'Shared TF',]


# maleTFORAList <- apply(CHOICE, 1, function(x) TF_ORA(geneset = TF_male_geneset,
#                                                      regulons=regulons,
#                                                          NESname = 'NES.male',
#                                                          PVname = NULL,
#                                                          genename = 'TF',
#                                                          cases = x, 
#                                                          bg = bg))
# names(maleTFORAList) <- listname
# 
# femaleTFORAList <- apply(CHOICE, 1, function(x) TF_ORA(geneset = TF_female_geneset,
#                                                      regulons=regulons,
#                                                      NESname = 'NES.female',
#                                                      PVname = NULL,
#                                                      genename = 'TF',
#                                                      cases = x, 
#                                                      bg = bg))
# names(femaleTFORAList) <- listname
# 
# sharedTFORAList <- apply(CHOICE, 1, function(x) TF_ORA(geneset = TF_shared_geneset,
#                                                        regulons=regulons,
#                                                        NESname = 'NES.female',
#                                                        PVname = NULL,
#                                                        genename = 'TF',
#                                                        cases = x, 
#                                                        bg = bg))
# names(sharedTFORAList) <- listname


# parallel
cl <- makeCluster (4)
clusterEvalQ(cl, {source('./fun/multiORA.R')
                  source('./fun/viper_function.R')})
clusterExport(cl, list("TF_male_geneset", "TF_female_geneset", "TF_shared_geneset","regulons","bg"))

maleTFORAList <- parApply(cl,  X = CHOICE, MARGIN = 1, FUN = function(x) TF_ORA(geneset = TF_male_geneset,
                                                     regulons=regulons,
                                                         NESname = 'NES.male',
                                                         PVname = NULL,
                                                         genename = 'TF',
                                                         cases = x, 
                                                         bg = bg))
names(maleTFORAList) <- listname

femaleTFORAList <- parApply(cl,  X = CHOICE, MARGIN = 1, FUN = function(x) TF_ORA(geneset = TF_female_geneset,
                                                     regulons=regulons,
                                                         NESname = 'NES.female',
                                                         PVname = NULL,
                                                         genename = 'TF',
                                                         cases = x, 
                                                         bg = bg))
names(femaleTFORAList) <- listname

sharedTFORAList <- parApply(cl,  X = CHOICE, MARGIN = 1, FUN = function(x) TF_ORA(geneset = TF_shared_geneset,
                                                       regulons=regulons,
                                                       NESname = 'NES.female',
                                                       PVname = NULL,
                                                       genename = 'TF',
                                                       cases = x, 
                                                       bg = bg))
names(sharedTFORAList) <- listname
stopCluster (cl)



```



# ORA VISUALIZATION
```{r}
# ORA VISUALIZATION
# Combine males and females results
PathwayLIST=list(maleTFORAList, femaleTFORAList)
names(PathwayLIST) = c('Male','Female')
PathwayAll_df <- reorganisedf(PathwayLIST)


showCompEnrichDotplot(PathwayAll_df,
                      pathwayname = 'GO',
                      split = TRUE,
                      top = 10,
                      wid = 1500,
                      height = 1000,
                      pcutoff = 0.05,
                      break_by = 5,
                      filename = "./results/ORATFGO top10 samplePerm.jpg")


showCompEnrichDotplot(PathwayAll_df,
                      pathwayname = 'pathway',
                      split = TRUE,
                      top = 10,
                      wid = 2200,
                      height = 1200,
                      pcutoff = 0.05,
                      break_by = 5,
                      filename = "./results/ORATFPathway top10 samplePerm.jpg")

```






Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
