---
title: "05_Pipeline_CARNIVAL"
output: html_notebook
---



```{r}
library(CARNIVAL)
library(OmnipathR)
source('./fun/CARNIVAL_functions.R')
library(progeny)
library(ggplot2)
library(ComplexHeatmap)

```


# run carnival
```{r}
progScorelist_SamPerm = readRDS('./data/progeny4carnival_sampleperm.rds')
TFactivity_Carnival_sampPerm = readRDS('./data/TF4Carnival_samplePerm.rds')


# progeny for CARNIVAL
load(file = system.file("progenyMembers.RData",package="CARNIVAL"))
omniR <- OmnipathR::import_Omnipath_Interactions()
sif = getSIF(omniR)


# get initial nodes
iniMTX = base::setdiff(sif$source, sif$target)
iniciators = base::data.frame(base::matrix(data = NaN, nrow = 1, ncol = length(iniMTX)), stringsAsFactors = F)
colnames(iniciators) = iniMTX


CHOICE = expand.grid(sex = c('male','female'),
                     p = c('normal','remo_unstable'),
                     perm = 'samplePerm')
CHOICE = sapply(CHOICE,as.character)
listname <- apply(CHOICE, 1 , function(x) paste(x, collapse = "_"))


adjpv = 3 # select adjpv = 3 
toplist = seq(from = 1, to=4, by = 0.1)
idx = which(toplist == adjpv)
CARNIVAL_results_ALL = apply(CHOICE, 1, function(x) CARNIVAL_Pipeline(x, idx,sif))
names(CARNIVAL_results_ALL) <- listname


```



# draw carnival network on cytoscape
```{r}
style.name = "CL1"
cytopath = './results/'

i = 7
carnival_result = CARNIVAL_results_ALL[[i]]
filename = names(CARNIVAL_results_ALL)[i]
draw_CYTOSCAPE_CARNIVAL(evis = carnival_result$weightedSIF,
                        nvis = carnival_result$nodesAttributes,
                        style.name = style.name,
                        title = filename)


saveSession(paste0(cytopath,filename)) #.cys
exportImage(paste0(cytopath,filename), 'PNG', zoom=200) #.png scaled by 200%
# exportImage(full.path, 'PDF') #.pdf


attrib <- carnival_result$attributesAll
sifall <- carnival_result$sifAll

sign = list()
edges = list()
for (i in 1:100) {
  x = attrib[[i]]
  y = sifall[[i]]
  if(any(x[,1] %in% 'RUNX2')){ 
    sign[[i]] = x[x[,1] %in% 'RUNX2', 2]
    edges[[i]] = y[y[,3] %in% 'RUNX2',]
} 
}

```



#ORA
```{r}
CHOICE = expand.grid(sex = c('male','female'),
                     p = c('normal','remo_unstable'),
                     perm = c('GenePerm','samplePerm'))

i=4
CARNICAL_male_node = CARNIVAL_results_ALL[[2*i-1]]$nodesAttributes
CARNICAL_female_node = CARNIVAL_results_ALL[[2*i]]$nodesAttributes
filename = paste(CHOICE[2*i-1,2],CHOICE[2*i-1,3]) 





CHOICEora = expand.grid(reg = c('up','dn'),
                     database = c('pathway','GO'))
CHOICEora = sapply(CHOICEora,as.character)
listnameora <- apply(CHOICEora, 1 , function(x) paste(x, collapse = "_"))


malePathwayList <- apply(CHOICEora, 1, function(x) multiORA(geneset = CARNICAL_male_node,
                                                        NESname = 'AvgAct',
                                                        PVname = NULL,
                                                        genename = 'Node',
                                                        x, 
                                                        bg = CARNICAL_male_node$Node,
                                                        pvcutoff = 0.05))

names(malePathwayList) <- listnameora
femalePathwayList <- apply(CHOICEora, 1, function(x) multiORA(geneset = CARNICAL_female_node,
                                                        NESname = 'AvgAct',
                                                        PVname = NULL,
                                                        genename = 'Node',
                                                        x, 
                                                        bg = CARNICAL_male_node$Node,
                                                        pvcutoff = 0.05))
names(femalePathwayList) <- listnameora

PathwayAll_df <- reorganisedf(femalePathwayList, malePathwayList)
showCompEnrichDotplot(PathwayAll_df,
                      pathwayname = 'GO',
                      split = TRUE,
                      top = 10,
                      pcutoff = 0.05,
                      wid = 1600,
                      height = 800,
                      filename = paste0("./results/limma+batch+age/CVD_vs_Control/carnivalGO_", filename,"_pcut0.05.jpg"))

showCompEnrichDotplot(PathwayAll_df,
                      pathwayname = 'pathway',
                      split = TRUE,
                      top = 10,
                      pcutoff = 0.05,
                      wid = 1500,
                      height = 800,
                      filename = paste0("./results/limma+batch+age/CVD_vs_Control/carnivalPathway_", filename,"_pcut0.05.jpg"))



```





```{r}
tfs = unique(regulons$tf)
tars = unique(regulons$target)



i = 4
carnival_result_ma=CARNIVAL_results_ALL[[2*i-1]]
carnival_result_fe=CARNIVAL_results_ALL[[2*i]]
filename = paste(CHOICE[2*i-1,2],CHOICE[2*i-1,3]) 


carnivalDegree_ma <- getCARNIVALdegree(carnival_result_ma$weightedSIF)
carnivalDegree_fe <- getCARNIVALdegree(carnival_result_fe$weightedSIF)
carnivalDegree <- carnivalDegree_fe %>%
  full_join(carnivalDegree_ma,by = "Node", suffix = c(".female", ".male"))


carnivalnodes_all = carnival_result_fe$nodesAttributes %>%
  inner_join(carnival_result_ma$nodesAttributes,by = "Node",
             suffix = c(".female", ".male")) %>%
  left_join(carnivalDegree, by = "Node" )

dtfrm <- carnivalnodes_all %>%
  mutate(diff= abs(AvgAct.female-AvgAct.male)) %>%
  arrange(desc(diff)) %>%
  # dplyr::select(Node, AvgAct.female, AvgAct.male, p.female, total_count.female, p.male, total_count.male) %>%
  column_to_rownames('Node')

dtfrm$TF <- NA
dtfrm$TF[rownames(dtfrm) %in% tars] <- 'Target'
dtfrm$TF[rownames(dtfrm) %in% tfs] <- 'TF'


top = 100
topdtfrm <- t(dtfrm[1:top,c("AvgAct.female","AvgAct.male")])
topdegree <- dtfrm[1:top,c("total_count.female","total_count.male")]
topdegree[is.na(topdegree)] <- 0

topanno <-  dtfrm[1:top,c("NodeType.female","NodeType.male",'TF')]

column_ha = HeatmapAnnotation( 
                              male = topanno$NodeType.male,
                              degree_ma = anno_text(topdegree$total_count.male, location = 0.5, rot = 0, just = "center"),
                              female = topanno$NodeType.female,
                              degree_fe = anno_text(topdegree$total_count.female, location = 0.5, rot = 0, just = "center"),
                              TF = topanno$TF,
                              col = list(male = c("T" = "pink",'S'='yellow'),
                              female = c("T" = "pink",'S'='yellow'),
                              TF = c("TF" = "pink", "Target" = "lightblue")),
                              annotation_name_side = "left")

pdfname <- paste0('./results/carnival_top',top,  filename,' heatmap.pdf')
pdf(pdfname, width = 22, height = 5)
ht_list = Heatmap(
  topdtfrm,
  top_annotation = column_ha,
  width = unit(ncol(topdtfrm)/2.5, "cm"),
  cluster_rows = FALSE,
  heatmap_legend_param = list(title = 'Activity Score')
)
draw(
  ht_list,
  # heatmap_legend_side = "left",
  heatmap_height = unit(nrow(topdtfrm)*3.5, "cm")
                                )
dev.off()


```





TOPology analysis of CARNIVAL
```{r}
# source('support_networks.r')
# source('CARNIVALnetwork.R')
# 
# 
# # get only summary files from CARNIVAL results
# sifts = list(male = carnival_result_ma$weightedSIF,female = carnival_result_fe$weightedSIF)
# nodos = list(male = carnival_result_ma$nodesAttributes,female = carnival_result_fe$nodesAttributes)
# interactions = getTopology(networks = sifts, scafoldNET = sif)

```











```{r}
# source('CARNIVALnetwork.R')
# 
# # saveCarnival <- function(PathwayActivity_CARNIVALinput,tf_activities_carnival){
# # carnival_result <- getCARNIVALresult(PathwayActivity_CARNIVALinput, tf_activities_carnival)
# # visNet = carnival_visNet(evis = carnival_result$weightedSIF,
# #                          nvis = carnival_result$nodesAttributes)
# # visSave(visNet, file = paste0('./results/limma+batch+age/CVD_vs_Control/',name,'_Prog=',top,'_carnival_visNetwork496_',flagname,'.html'), selfcontained = TRUE)
# # save(visNet, carnival_result, file = paste0('./results/limma+batch+age/CVD_vs_Control/',name,'_Prog=',top,'_carnival_visNetwork496_',flagname,'.RData'))
# # }
# 
# # male carnival
# name = "male"
# PathwayActivity_CARNIVALinput = progenyScore %>%
#   tibble::column_to_rownames(var = 'Pathway')%>%
#   dplyr::select(male)
# tf_activities_carnival <- TF_activity_all %>%
#   tibble::column_to_rownames(var = 'TF')%>%
#   filter(labels == 'Male specific TF') %>%
#    dplyr::select(NES.male)
# 
# # saveCarnival(PathwayActivity_CARNIVALinput,tf_activities_carnival)
# carnival_result <- getCARNIVALresult(PathwayActivity_CARNIVALinput, tf_activities_carnival)
# visNet = carnival_visNet(evis = carnival_result$weightedSIF,
#                          nvis = carnival_result$nodesAttributes)
# visSave(visNet, file = paste0('./results/limma+batch+age/CVD_vs_Control/',name,'_Prog=',top,'_carnival_visNetwork496_',flagname,'.html'), selfcontained = TRUE)
# 
# 
# save(visNet, carnival_result, file = paste0('./results/limma+batch+age/CVD_vs_Control/',name,'_Prog=',top,'_carnival_visNetwork496_',flagname,'.RData'))


    
```



















TOPology analysis of CARNIVAL
```{r}
# source('support_networks.r')
# source('CARNIVALnetwork.R')
# 
# load("./results/limma+batch+age/CVD_vs_Control/female_tfpv=0.1_logprogpv=6_carnival_result496.RData")
# carnival_result_fe = carnival_result
# load("./results/limma+batch+age/CVD_vs_Control/male_tfpv=0.1_logprogpv=6_carnival_result496.RData")
# carnival_result_ma = carnival_result
# 
# sif = getSIF()
# # get only summary files from CARNIVAL results
# sifts = list(male = carnival_result_ma$weightedSIF,female = carnival_result_fe$weightedSIF)
# nodos = list(male = carnival_result_ma$nodesAttributes,female = carnival_result_fe$nodesAttributes)
# interactions = getTopology(networks = sifts, scafoldNET = sif)
# 
# carnivalDegree_ma <- getCARNIVALdegree(carnival_result_ma$weightedSIF)
# carnivalDegree_fe <- getCARNIVALdegree(carnival_result_fe$weightedSIF)
# carnivalDegree <- carnivalDegree_fe %>%
#   full_join(carnivalDegree_ma,by = "Node", suffix = c(".female", ".male"))
# 
# 
# carnivalnodes_all = carnival_result_fe$nodesAttributes %>%
#   inner_join(carnival_result_ma$nodesAttributes,by = "Node",
#              suffix = c(".female", ".male")) %>%
#   left_join(carnivalDegree, by = "Node" )
```





```{r}
# library(ggplot2)
# library(ComplexHeatmap)
# 
# tfs = unique(regulons$tf)
# tars = unique(regulons$target)
# 
# dtfrm <- carnivalnodes_all %>%
#   mutate(diff= abs(AvgAct.female-AvgAct.male)) %>%
#   arrange(desc(diff)) %>%
#   # dplyr::select(Node, AvgAct.female, AvgAct.male, p.female, total_count.female, p.male, total_count.male) %>%
#   column_to_rownames('Node')
# 
# dtfrm$TF <- NA
# dtfrm$TF[rownames(dtfrm) %in% tars] <- 'Target'
# dtfrm$TF[rownames(dtfrm) %in% tfs] <- 'TF'
# 
# 
# top = 100
# topdtfrm <- t(dtfrm[1:top,c("AvgAct.female","AvgAct.male")])
# topdegree <- dtfrm[1:top,c("total_count.female","total_count.male")]
# topdegree[is.na(topdegree)] <- 0
# 
# topanno <-  dtfrm[1:top,c("NodeType.female","NodeType.male",'TF')]
# 
# column_ha = HeatmapAnnotation( 
#                               male = topanno$NodeType.male,
#                               degree_ma = anno_text(topdegree$total_count.male, location = 0.5, rot = 0, just = "center"),
#                               female = topanno$NodeType.female,
#                               degree_fe = anno_text(topdegree$total_count.female, location = 0.5, rot = 0, just = "center"),
#                               TF = topanno$TF,
#                               col = list(male = c("T" = "pink", "S" = "yellow"),
#                               female = c("T" = "pink", "S" = "yellow"),
#                               TF = c("TF" = "pink", "Target" = "lightblue")),
#                               annotation_name_side = "left")
# pdfname <- paste0('./results/limma+batch+age/CVD_vs_Control/tfpv=0.1_logprogpv=6_carnival_top',top,' 496.pdf')
# pdf(pdfname, width = 22, height = 5)
# ht_list = Heatmap(
#   topdtfrm,
#   top_annotation = column_ha,
#   width = unit(ncol(topdtfrm)/2.5, "cm"),
#   cluster_rows = FALSE,
#   heatmap_legend_param = list(title = 'Activity Score')
# )
# draw(
#   ht_list,
#   # heatmap_legend_side = "left",
#   heatmap_height = unit(nrow(topdtfrm)*3.5, "cm")
#                                 )
# dev.off()


```







Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
